Bayesian Sparsification of $\cplx$-valued networks

### 2020-01-30 paper/paper.tex
edits and plans

In this section we compare the complex variational dropout techniques discussed above
on the tasks and datasets, that were studied in prior research on complex-valued networks.
All experiments share the following settings unless specified otherwise:
\begin{itemize}
  \item ADAM optimizer with the base learning rate of $10^{-3}$
  \item learning rate the annealing schedule from \citep{trabelsi_deep_2017}
  \item global gradient clipping to within an $\ell_2$-norm ball of radius $0.05$
\end{itemize}

@@

\begin{figure}[!ht]
  \centering
  \begin{subfigure}[b]{0.9\textwidth}
    \centering
    \includegraphics[width=1.\linewidth,draft]{../notebooks/assets/grad_log.pdf}
    % \caption{} \label{fig:log-relevance-kde}
  \end{subfigure} \\
  % \hfill
  \begin{subfigure}[b]{0.9\textwidth}
    \centering
    \includegraphics[width=1.\linewidth,draft]{../notebooks/assets/grad_log.pdf}
    % \caption{} \label{fig:log-relevance-proba}
  \end{subfigure}
  \caption{%
    \textit{top}: Estimate of $\log\alpha$ density for a convolutional and dense layer
    from the convolutional model of sec.~\ref{sub:mnist_like_datasets}.
    \textit{bottom}: The probability of 
  }
  \label{fig:log-relevance}
\end{figure}

We decide if a parameter is \textit{``non-zero''} if it is within $\delta$ relative
tolerance of its mean with high probability under the approximate posterior. Since
\eqref{eq:c-gauss-vi} is Gaussian, $
  \tfrac{\lvert w - \mu \rvert^2}
        {\alpha \lvert \mu \rvert^2}
$ is $\chi^2_k$ distributed with shape $k=1$ ($\real$) or $2$ ($\cplx$), whence
\begin{equation} \label{eq:prob_relevant}
  \Pr\bigl(
    \lvert w - \mu \rvert \leq \delta \lvert \mu \rvert
  \bigr)
    = \Pr\biggl(
      \frac{\lvert w - \mu \rvert^2}{\alpha \lvert \mu \rvert^2}
        \leq \frac{\delta^2}{\alpha}
    \biggr)
    = \chi^2_k\biggl(\frac{\delta^2}{\alpha}\biggr)
  \,.
\end{equation}
Figure~\ref{fig:log-relevance} (bottom) shows the probability of a $\real$ (or $\cplx$) 
parameter being non-zero for various $\delta\in(0, \tfrac12)$ as a function of $\log \alpha$.
For a reasonably generous $\delta = 25\%$ tolerance all values of $\log \alpha$ below
$-2.5$ yield a reasonably high chance of a non-zero parameter. Given that $q_\theta$ is
merely an approximation we opt for a slightly higher $\log \alpha$ threshold, and thus
set $\tau = -\tfrac12$ in every experiment, despite it being lower than the recommended
$\tau = 3$ in \citep{molchanov_variational_2017}, but yields much higher levels of sparsity.
The rationale behind our choice was twofold: we use an extra fine tuning stage, and keep the
parameters that are away from zero with high probability.

@@

We conduct a moderately sized numerical experiment on MNIST and similar datasets to see
whether the proposed $\cplx$-var dropout indeed reasonably sparsifies complex-valued networks.

We compare with the studied $\cplx$-valued networks against their classical $\real$-valued
counterparts both in terms of the resulting sparsity and test performance.

Although image data is not $\cplx$ in nature, we apply two-dimensional Fourier Transform
in order to get $\cplx$-valued inputs. We concede that this is a rather unnatural representation
of image data, however the selection of publicly available datasets, which are not artificially
$\cplx$-valued, is limited to audio, speech and music. Indeed, as is shown by the numerical
experiments, a real-valued model is trained on the Fourier transformed data underperforms
relative to its architecturally similar twin fit to on the raw color data.

\citet{trabelsi_deep_2017} study the applicability of $\cplx$-valued networks to image
classification (CIFAR-10, CIFAR-100, and a subset of SVHN). Their approach is to either
embed $\real$ in $\cplx$ (by $\Re z = 0$), or train a smaller auxiliary network, that
synthesizes the imaginary component from the raw channel data. \todo{DO THIS}
In our experiments we opt to transforming the input $2d$ data with

>>>>

In this section we study the complex variational dropout techniques discussed above
on the tasks and datasets, that were studied in prior research on complex-valued networks.
All experiments share the following settings unless specified otherwise:
\begin{itemize}
  \item ADAM optimizer with the base learning rate of $10^{-3}$
  \item learning rate the annealing schedule from \citep{trabelsi_deep_2017}
  \item global gradient clipping within an $\ell_2$-norm ball of radius $0.05$
\end{itemize}

@@

\begin{figure}[!ht]
  \centering
  \begin{subfigure}[b]{0.9\textwidth}
    \centering
    \includegraphics[width=1.\linewidth,draft]{../notebooks/assets/log-alpha__histograms.pdf}
    % \caption{} \label{fig:log-relevance-kde}
  \end{subfigure} \\
  % \hfill
  \begin{subfigure}[b]{0.9\textwidth}
    \centering
    \includegraphics[width=1.\linewidth,draft]{../notebooks/assets/proba_of_nonzero.pdf}
    % \caption{} \label{fig:log-relevance-proba}
  \end{subfigure}
  \caption{%
    \textit{top}: Estimate of $\log\alpha$ density for a convolutional and dense layer
    from the convolutional model of sec.~\ref{sub:mnist_like_datasets}.
    \textit{bottom}: The probability of a parameter being within $\pm \delta$ relative
    tolerance of its mean.
  }
  \label{fig:log-relevance}
\end{figure}

We decide if a parameter is \textit{``non-zero''} if it is within $\delta$ relative
tolerance of its mean with high probability under the approximate posterior. Since
\eqref{eq:c-gauss-vi} is Gaussian, $
  \tfrac{\lvert w - \mu \rvert^2}
        {\alpha \lvert \mu \rvert^2}
$ is $\chi^2_k$ distributed with shape $k=1$ ($\real$) or $2$ ($\cplx$), whence
\begin{equation} \label{eq:prob_relevant}
  \Pr\bigl(
    \lvert w - \mu \rvert \leq \delta \lvert \mu \rvert
  \bigr)
    = \Pr\biggl(
      \frac{\lvert w - \mu \rvert^2}{\alpha \lvert \mu \rvert^2}
        \leq \frac{\delta^2}{\alpha}
    \biggr)
    = \chi^2_k\biggl(\frac{\delta^2}{\alpha}\biggr)
  \,.
\end{equation}
Figure~\ref{fig:log-relevance} (bottom) shows the probability of a $\real$ (or $\cplx$) 
parameter being non-zero for various $\delta\in(0, \tfrac12)$ as a function of $\log \alpha$.
For a reasonably generous $\delta = 25\%$ tolerance all values of $\log \alpha$ below
$-2.5$ yield a relatively high chance of a non-zero parameter. Given that $q_\theta$
is merely an approximation, we opt for a somewhat higher $\log \alpha$ threshold, and
pick $\tau = -\tfrac12$, which is much lower than $\tau = 3$, recommended by
\citet{molchanov_variational_2017}. Additional rationale behind this decision is the
extra fine tuning, which trains the compressed model.

@@

We conduct a moderately sized numerical experiment on MNIST and similar $28\times 28$
greyscale image datasets to see whether the proposed $\cplx$-variational dropout indeed
sparsifies complex-valued networks. We compare with the studied $\cplx$-valued networks
against their $\real$-valued counterparts both in terms of the resulting sparsity and
test performance. For a fair comparison, we make the $\cplx$-valued network have
approximately half as many parameters, \citep{monning_evaluation_2018}.

Although image data is not $\cplx$ in nature \citet{trabelsi_deep_2017} study the applicability
of $\cplx$-valued networks to image classification (CIFAR-10, CIFAR-100, and a subset of
SVHN). Their approach is to either embed $\real$ in $\cplx$ (by $\Re z = 0$), or train
a smaller auxiliary network, that synthesizes the imaginary component from the raw channel
data. In our experiments we apply the two-dimensional Fourier Transform to get $\cplx$-valued
inputs. We concede that this is a rather unnatural representation of image data, however
the selection of publicly available datasets, which are not artificially $\cplx$-valued,
is limited to audio, speech and music.

Indeed, as is shown by the numerical experiments on MNIST-like datasets and CIFAR10, a
real-valued model trained on the Fourier transformed data underperforms relative to
its architectural twin, fit to on the raw color data. At the same time complex-valued
models

% a copule of plot form mnist and fashion mnist
% discuss the plots describing how the test -perfromance grows with compression
% hinting at overfitting/regularization effects of masking and sparsification
